---
title: "Chicago Airbnb Analysis"
author: "Jose Luis Rodriguez"
date: "June 13, 2018"
output:
  html_notebook: default
subtitle: CME Group Foundation Business Analytics Lab
---

-------------

## Notebook Instructions

-------------

* This notebook is use to analyze airbnb data based on the Inside Airbnb project by Murray Cox.

#### Notebook Version

```
platform       x86_64-apple-darwin15.6.0   
arch           x86_64                      
os             darwin15.6.0                
system         x86_64, darwin15.6.0        
status                                     
major          3                           
minor          5.0                         
year           2018                        
month          04                          
day            23                          
svn rev        74626                       
language       R                           
version.string R version 3.5.0 (2018-04-23)
nickname       Joy in Playing   
```

### About

* Airbnb claims to be part of the "sharing economy" and disrupting the hotel industry. 

* However, data shows that the majority of Airbnb listings in most cities are entire homes, many of which are rented all year round - disrupting housing and communities.

* http://insideairbnb.com/index.html

### Project Layout

```
.
└── Airbnb
    ├── airbnb-notebook.Rmd
    │
    ├── data
    │   ├── 2015
    │   │   ├── airbnb.csv
    │   │   ├── airbnb_calendar.csv
    │   │   ├── airbnb_clean.csv
    │   │   ├── airbnb_loc.csv
    │   │   ├── airbnb_nlp.csv
    │   │   ├── airbnb_reviews.csv
    │   │   ├── airbnb_summary.csv
    │   │   └── airbnb_urls.csv
    │   │
    │   ├── 2017
    │   │   ├── airbnb.csv
    │   │   ├── airbnb_calendar.csv
    │   │   ├── airbnb_reviews.csv
    │   │   └── airbnb_summary.csv
    │   │
    │   ├── location
    │   │   ├── 2015-neighbourhoods.csv
    │   │   ├── 2015-neighbourhoods.geojson
    │   │   ├── 2017-neighbourhoods.csv
    │   │   ├── 2017-neighbourhoods.geojson
    │   │   └── names-zipcode.csv
    │   │
    │   └── raw
    │       ├── 2015-airbnb-listings.csv.gz
    │       ├── 2015-calendar.csv.gz
    │       ├── 2015-neighbourhoods.zip
    │       ├── 2015-reviews.csv.gz
    │       ├── 2017-airbnb-listings.csv.gz
    │       ├── 2017-calendar.csv.gz
    │       ├── 2017-neighbourhoods.zip
    │       └── 2017-reviews.csv.gz
    │
    ├── resources
    │   ├── Airbnb-Proposal.docx
    │   └── Airbnb-Proposal.pdf
    │
    └── scripts
```


### Load Packages in R/RStudio 

We are going to use tidyverse a collection of R packages designed for data science. 

* Info: https://www.tidyverse.org

```{r, echo = FALSE}

# Here we are checking if the package is installed
if(!require("tidyverse")){
  
  # If the package is not in the system then it will be install
  install.packages("tidyverse", dependencies = TRUE)
  
  # Here we are loading the package
  library("tidyverse")
}

```

-------------

## Data Import and Inspection

-------------

* read dataset from 2015

```{r}

airbnb <- read_csv("data/2015/airbnb.csv")

```

### Inspect head and tail of the dataset
```{r}

head(airbnb)

```

```{r}

tail(airbnb)

```

```{r}

colnames(airbnb)

```

-------------

## Data Selection

-------------

* Now lets clean the data and delete some unessary columns for this analysis

* Extract the airbnb_id to later use and identify the columns to remove

* Since there are 92 features/columns we can split the column names vector into different chunks 

* The 92 features/columns are of variating types and measures we can split the dataset into diffent sections. 

* Such as features/columns related to links/urls or only text columns that we can use later for natural language processing.

* Same process applies to features/columns related to location 

* Finally we can remove the columns from the new datasets to create a final dataset that will be more numeric in nature. 

* All datasets will be linked/related by the airbnb_id


```{r}
airbnb <- rename(airbnb, airbnb_id = id)
airbnb_id <- airbnb$airbnb_id
```

```{r}
colnames(airbnb)[1:30]
```

```{r}
colnames(airbnb)[31:65]
```

```{r}
colnames(airbnb)[66:92]
```

### Select only columns related to url links

```{r}
airbnb_urls <- select(airbnb, contains("url"))
ulr_columns <- colnames(airbnb_urls)
airbnb_urls <- add_column(airbnb_urls, airbnb_id)

```

```{r}
head(airbnb_urls)
```

### Select only columns containg text 


```{r}
text_columns <- c("host_name","name","amenities", "experiences_offered",
                  "host_verifications","transit","notes", 
                 "neighborhood_overview", "host_about", 
                 "description", "space","summary")

airbnb_nlp <- select(airbnb, text_columns)

airbnb_nlp <- add_column(airbnb_nlp, airbnb_id)

```

```{r}
head(airbnb_nlp)
```

### Select only columns related to listing location

```{r}

location_columns <- c("zipcode", "latitude", "longitude", "city", "state",
                     "street", "neighbourhood", "neighbourhood_cleansed",
                     "neighbourhood_group_cleansed","host_location", "host_neighbourhood",
                     "market", "smart_location", "jurisdiction_names", "country_code", 
                     "country")

airbnb_loc <- select(airbnb, location_columns)

airbnb_loc <- add_column(airbnb_loc, airbnb_id)

```

```{r}
head(airbnb_loc)
```

### Use select and one_of function to remove the columns from previous datasets (ulr_columns, text_columns, location_columns)

```{r}
# unique(as.character(airbnb$scrape_id))

remove_columns <- c(ulr_columns, text_columns, location_columns, "scrape_id", "last_scraped")

airbnb_clean <- select(airbnb, -one_of(remove_columns))

```

### Inspect the new more compacted dataset. By doing this we reduced the size by almost half

```{r}

head(airbnb_clean)

```

### Save the new dataset for reference and later use

```{r}
write_csv(airbnb_urls, "data/2015/airbnb_urls.csv")

write_csv(airbnb_nlp, "data/2015/airbnb_nlp.csv")

write_csv(airbnb_loc, "data/2015/airbnb_loc.csv")

write_csv(airbnb_clean, "data/2015/airbnb_clean.csv")

```

-------------

## Data Types and Cleaning

-------------


```{r}
airbnb_clean <- read_csv("data/2015/airbnb_clean.csv")
head(airbnb_clean)
```


```{r}
room_type <- airbnb_clean$room_type %>% 
  tolower() %>% 
  str_replace_all(., " ", "_") %>% 
  str_replace(., "home/apt","place")

airbnb_clean$room_type <- as_factor(room_type)

```


```{r}

# unique(airbnb_clean$bed_type)
bed_type <- airbnb_clean$bed_type %>% 
  tolower() %>% 
  str_replace_all(., " ", "_") %>% 
  str_replace(., "pull.*","sofa_bed") %>% 
  str_replace(., "real_.*","bed")

airbnb_clean$bed_type <- as_factor(bed_type)

```


```{r}

airbnb_clean <- airbnb_clean %>% 
  mutate(airbnb_id = as.character(airbnb_id)) %>% 
  mutate(host_id = as.character(host_id)) %>% 
  mutate(host_is_superhost = parse_logical(host_is_superhost)) %>% 
  mutate(host_has_profile_pic = parse_logical(host_has_profile_pic)) %>% 
  mutate(host_identity_verified = parse_logical(host_identity_verified)) %>% 
  mutate(is_location_exact = parse_logical(is_location_exact)) %>% 
  mutate(property_type = as_factor(tolower(property_type)))

```

```{r}
head(airbnb_clean)
```

```{r}
summary(airbnb_clean)
```


```{r}

host_response_rate <- airbnb_price$host_response_rate
host_response_rate <- as.numeric(str_remove(host_response_rate, "%"))
host_response_rate <- host_response_rate/100
airbnb_price$host_response_rate <- host_response_rate

```

```{r}

host_verifications <- str_remove(airbnb_price$host_verifications[1], "\\[")
host_verifications <- str_remove(host_verifications, "\\]")
host_verifications <- str_remove_all(host_verifications, "\'")
host_verifications <- str_split(host_verifications, ",") 

host_verifications

```

### Save clean dataset with different name

```{r}

write_csv(airbnb_price, "data/airbnb_price.csv")

```

-------------

## Descriptive Statistics 

-------------

```{r}
summary(airbnb_price[1:10])
```

```{r}
summary(airbnb_price[10:20])
```
