---
title: "Chicago Airbnb Analysis"
author: "Jose Luis Rodriguez"
date: "June 13, 2018"
output:
  html_notebook: default
subtitle: CME Group Foundation Business Analytics Lab
---

-------------

## Notebook Instructions

-------------

* This notebook is use to analyze airbnb data based on the Inside Airbnb project by Murray Cox.

### About

* Airbnb claims to be part of the "sharing economy" and disrupting the hotel industry. 

* However, data shows that the majority of Airbnb listings in most cities are entire homes, many of which are rented all year round - disrupting housing and communities. 


* http://insideairbnb.com/index.html

### Load Packages in R/RStudio 

We are going to use tidyverse a collection of R packages designed for data science. 

* Info: https://www.tidyverse.org/

```{r, echo = FALSE}

# Here we are checking if the package is installed
if(!require("tidyverse")){
  
  # If the package is not in the system then it will be install
  install.packages("tidyverse", dependencies = TRUE)
  
  # Here we are loading the package
  library("tidyverse")
}

```

### Read the data from 2015

```{r}

listings <- read_csv("data/2017/listings.csv")

listings
```

```{r}

```


